---
title:  Livestock Breeding and Genomics - Exercise 2
author: Peter von Rohr
date: 2021-10-15
output: html_notebook
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```


## Problem 1: Breeding Values For a Monogenic Trait
We assume that the absorption of cholesterol is determined by a certain enzyme. The level of enzyme production is determined by a single bi-allelic locus $E$. The genotype frequencies and the genotypic values for the two dairy cattle populations `Original Braunvieh` and `Brown Swiss` are given in the following table.

```{r table-geno-freq-value, echo=FALSE}
l_maf <- list(ob = 0.25, bs = 0.1)
l_a <- list(ob = 15, bs = 29)
l_d <- list(ob = 3, bs = 0)
tbl_geno_freq_value <- tibble::tibble(Variable = c(
  "$f(E_1E_1)$",
  "$f(E_1E_2)$",
  "$f(E_2E_2)$",
  "$a$",
  "$d$"),
  `Original Braunvieh` = c(
    l_maf$ob^2, 
    2*l_maf$ob*(1-l_maf$ob),
    (1-l_maf$ob)^2,
    l_a$ob,
    l_d$ob
  ),
  `Brown Swiss` = c(
    l_maf$bs^2,
    2*l_maf$bs * (1-l_maf$bs),
    (1-l_maf$bs),
    l_a$bs,
    l_d$bs
  ))
knitr::kable(tbl_geno_freq_value, booktabs = TRUE, escape = FALSE)
```


### Hints
* Assume that allele $E_1$ is the allele with the positive effect on the enzyme level
* Assume that the Hardy-Weinberg Equilibrium holds in both populations


### Your Task
Compute the breeding values for all three genotypes in both populations. 


### Your Solution


## Problem 2: Quantitative Genetics
In a population the following numbers of genotypes were counted for a given genetic locus called $A$. 

```{r GenotypeFreqs, echo=FALSE, results='asis'}
dfGenotypeFreq <- data.frame(Genotypes = c("$A_1A_1$", "$A_1A_2$", "$A_2A_2$"),
                             Numbers   = c(24, 53, 23),
                             stringsAsFactors = FALSE)
knitr::kable(dfGenotypeFreq)
```

a) Compute the genotype frequencies

b) Compute the allele frequencies

c) Compute the population mean $\mu$ under the following assumptions

* the difference between the genotypic values of the homozygous genotypes is $20$ and
* the genotypic value of the heterozygous genotype is $2$.

### Your Solution
