---
title: "Design and Development of Course Package Features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design and Development of Course Package Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(lbgfs2021)
# rmdhelp::show_knit_hook_call()
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```


# Disclaimer
Design ideas and potentially useful features for a course R-package are developed and described.


# General Design Idea
The general idea of encapsulating course material in R-packages is to create a hierarchy of packages. On the top-level there is a generic R-package that is used to initialise a child package which is specific for a given course. The following diagram shows the package hierarchy. 

```{r pkg-hierarchy}
rmdhelp::use_odg_graphic(ps_path = "odg/pkg-hierarchy.odg")
```



# Features

## Templates

