---
title: "Livestock Breeding and Genomics - Solution 3"
author: "Peter von Rohr"
date: "10/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Own Performance
```{r prob1-setup, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(4206)
n_nr_animal <- 12
n_weight_mean <- 283
n_weight_sd <- 3
vec_weight <- round(rnorm(n_nr_animal, mean = n_weight_mean, n_weight_sd), digits = 0)
tbl_weight <- tibble::tibble(Animal = c(1:n_nr_animal),
                             Weight = vec_weight)
n_weight_h2 <- 0.2025
```

Given is the dataset with weight observations for `r n_nr_animal` animals. The heritability ($h^2$) for the trait is `r n_weight_h2`. The population mean $\mu$ can assumed to be the mean of the weights in the table below. 

`
```{r tbl-weight, echo=FALSE}
knitr::kable(tbl_weight)
```

### Your Tasks

* Compute the breeding values for all animals given in the table above
* Compute the accuracies of the breeding values of all animals shown in the table above.


### Solution
The predicted breeding value $\widehat{u_i}$ of animal $i$ is computed as 

$$\widehat{u_i} = h^2 (y_i - \mu)$$

where $h^2$ is the heritability given in the problem, $y_i$ is observation of animal $i$ and $\mu$ is the poputlation mean. The population mean is to be computed from the mean of the observations. Hence

```{r}
n_mu_weight <- mean(tbl_weight$Weight)
```




