---
title: "Livestock Breeding and Genomics - Excercise 7"
author: "Peter von Rohr"
date: '2021-11-19'
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pedigree

```{r}
nr_animal <- 6
tbl_pedigree <- tibble::tibble(Calf = c(1:nr_animal),
                               Sire = c(NA, NA, NA, 1 ,3, 4),
                               Dam = c(NA, NA, NA, 2, 2, 5))
tbl_pedigree
```


# Numerator Relationship Matrix
We call the numerator relationship matrix $A$. The computation of the elements of $A$ are done separately for 

1. the diagnoal elements $(A)_{ii}$ and
2. the off-diagonal elements $(A)_{ij}$ for $i \ne j$


First all elements of the matrix $A$ are initialized to $0$

```{r}
A = matrix(0, nrow = nr_animal, ncol = nr_animal)
A
```


## Diagonal Elements

Computation: $(A)_{ii} = (1 + F_i)$ and $F_i = 1/2 (A)_{sd}$ 

```{r}
i <- 1
s <- tbl_pedigree$Sire[i]
d <- tbl_pedigree$Dam[i]

Fi <- ifelse((is.na(s) | is.na(d)), 0, 0.5 * A[s,d])
A[i,i] <- 1+Fi
A

```


## Off-diagonal Elements

Off-diagonal $(A)_{ij} = 1/2 (A)_{io} + 1/2 (A)_{iq}$ where $o$ and $q$ are parents of $j$


```{r}
for (j in (i+1):6){
  o <- tbl_pedigree$Sire[j]
  q <- tbl_pedigree$Dam[j]
  Aio <- ifelse(is.na(o), 0, A[i,o])
  Aiq <- ifelse(is.na(q), 0, A[i,q])
  A[i,j] <- 0.5 * Aio + 0.5 * Aiq
}
A[(i+1):6,i] <- A[i,(i+1):6]
A
```


# Problem 1: Numerator Relationship Matrix
Use the above steps of computation for the complete matrix. 


## Hint

* Construct a loop with loop-variable $i$ that runs over all the rows of the matrix $A$.


## Solution


## Check Result
The function `getA()` of the pedigreemm package can be used to check the result















