---
title: Document Title
author: pvr
date: "`r Sys.Date()`"
output: output_format
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data
The same dataset is used as for the explanation of the Mixed Model Equation for the sire model. The data can be read from

```{r}
s_data_url <- "https://charlotte-ngs.github.io/lbgfs2021/data/beef_data_bc.csv"
tbl_beef_bc <- readr::read_csv(file = s_data_url)
tbl_beef_bc$Herd <- as.factor(tbl_beef_bc$Herd)
tbl_beef_bc$Sire <- as.factor(tbl_beef_bc$Sire)
head(tbl_beef_bc)
```


# Model

```{r}
library(lme4)
fit_beef_bc_all <- lmer(formula = `Weaning Weight` ~ Herd + `Breast Circumference` + (1|Sire), data = tbl_beef_bc)
isSingular(fit_beef_bc_all)
```

Reducing model complexity

```{r}
fit_beef_sire_no_inter <- lmer(formula = `Weaning Weight` ~  0 + (1|Sire), data = tbl_beef_bc)
isSingular(fit_beef_sire_no_inter)
```

Model summary

```{r}
summary(fit_beef_sire_no_inter)
```


# Data from lme4 Vignette
The package functionality is tested with a dataset that is documented to be working. 

```{r}
library(lme4)
str(sleepstudy)
```


Descriptive statistics reveals a linear dependence of `Reaction` on `Days` but each subject seams to have a different slope in the relationship. This is modelled with the following statement.

```{r}
fm1 <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
summary(fm1)
```

