---
title: Livestock Breeding and Genomics - Solution 11
author: Peter von Rohr
date: "2021-12-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

```{r}
nr_animal <- 5
tbl_decomp <- tibble::tibble(Animal = c(1:nr_animal),
                             Sire = c(NA, NA, NA, 1, 4),
                             Dam = c(NA, NA, NA, 2, 3),
                             Trait = c(4.5, 2.9, 3.9, 3.5, 5.0))
tbl_decomp
```


# Model

$$y = X\mu + Zu + e$$


# Components

```{r}
X = matrix(1, nrow = nr_animal, ncol = 1)
X
```

```{r}
Z = diag(1, nrow = nr_animal)
Z
```


# Pedigree

```{r}
ped = pedigreemm::pedigree(sire = tbl_decomp$Sire, 
                           dam = tbl_decomp$Dam,
                           label = as.character(1:nr_animal))
Ainv <- as.matrix(pedigreemm::getAInv(ped = ped))
Ainv
                  
```


# MME

```{r}
xtx <- crossprod(X)
xtx
```

```{r}
xtz <- crossprod(X, Z)
xtz
```

```{r}
ztx <- crossprod(Z, X)
ztx
```

```{r}
ztz <- crossprod(Z)
ztz
```

```{r}
sigmae2 <- 40
sigmau2 <- 20
lambda <- sigmae2/sigmau2
lambda
```

```{r}
ztzainvlambda <- ztz + Ainv * lambda
ztzainvlambda
```

```{r}
coef_mat <- rbind(cbind(xtx, xtz), cbind(ztx, ztzainvlambda))
coef_mat
```

```{r}
y <- tbl_decomp$Trait
xty <- crossprod(X,y)
zty <- crossprod(Z,y)
rhs <- rbind(xty, zty)
rhs 

```

# Solutions

```{r}
sol <- solve(coef_mat, rhs)
sol
```


# Problem 1: Decomposition
From the lecture, we know how the predicted breeding value of animal $4 can be decomposed into different components. Verify this decomposition based on the solutions obtained above.


## Solution
The decomposition is given as follows:

$$\hat{u}_4 = {1\over 6}\left[ y_4 - \hat{\mu} + 2 * (\hat{u}_1 + \hat{u}_2) - \hat{u}_3 + 2 \hat{u}_5\right] $$

This can be verified as follows

```{r}
(tbl_decomp$Trait[4] - sol[1] + 2 * (sol[2]+sol[3]) - sol[4] + 2*sol[6])/6
```

The breeding value of animal 4 is

```{r}
sol[5]
```


# Problem 2: Reliabilities
Compute the reliabilities of all the predicted breeding values. The reliability $B_i$ of the predicted breeding value for animal $i$ is computed as

$$B_i = 1 - \frac{(C^{22})_{ii}}{var(u_i)} $$

```{r}
C <- solve(coef_mat) * sigmae2
C
```

The lower right corner $C^{22}$ is 

```{r}
nr_fixed_effect <- ncol(X)
nr_sol <- nrow(sol)
C22 <- C[(nr_fixed_effect+1):nr_sol, (nr_fixed_effect+1):nr_sol]
C22
```

The diagonal elements of $C^{22}$

```{r}
C22ii <- diag(C22)
C22ii
```

The reliabilities are 

```{r}
B <- 1 - C22ii / sigmau2
B
```

