---
title: Data for Exercise 5
author: Peter von Rohr
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer
Based on beef dataset, an extended dataset is created to show the difference between fixed effects and covariates. 


# Read Existing Data
We start by reading the existing beef dataset

```{r}
s_data_path <- file.path(here::here(), "docs", "data", "beef_data_blup.csv")
tbl_beef_data <- readr::read_csv(file = s_data_path)
tbl_beef_data
```


# Additional Variable Breast Circumference
Breast Circumference of animals are added as an additional data column. It is computed from a previous regression. 

```{r}
set.seed(731)
# constants
n_inv_reg_coeff <- 1.7
n_bc_low_limit <- 1.32
n_bc_up_limit <- 1.78
n_nr_animal <- nrow(tbl_beef_data)
# compute bc based on inv reg
vec_bc_comp <- tbl_beef_data$`Weaning Weight` / n_inv_reg_coeff
```

Check summary

```{r}
summary(vec_bc_comp)
```

Histogram

```{r}
hist(vec_bc_comp)
```


# Adjustments
Do adjustments with lower and upper bound

```{r}
vec_bc_sim <- vec_bc_comp + rnorm(n_nr_animal, mean = 0, sd = 0.1)
vec_bc_too_small <- which(vec_bc_sim < n_bc_low_limit)
vec_bc_too_high <- which(vec_bc_sim > n_bc_up_limit)
vec_bc_sim[vec_bc_too_small] <- n_bc_low_limit + 
  abs(rnorm(length(vec_bc_too_small), mean = 0, sd = 0.2))
vec_bc_sim[vec_bc_too_high] <- n_bc_up_limit - 
  abs(rnorm(length(vec_bc_too_high), mean = 0, sd = 0.2))
```

Check

```{r}
summary(vec_bc_sim)
```

Histogram

```{r}
hist(vec_bc_sim)
```

# Adding Data

```{r}
tbl_beef_data_bc <- dplyr::bind_cols(tbl_beef_data, 
                                     tibble::tibble(`Breast Circumference` = round(vec_bc_sim, digits = 2)))
tbl_beef_data_bc
```

