---
title: Livestock Breeding and Genomics - Solution 4
author: Peter von Rohr
date: "2021-10-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r prob1-init, echo=FALSE, message=FALSE, warning=FALSE}
s_data_url <- "https://charlotte-ngs.github.io/lbgfs2021/data/beef_data_blup.csv"
```

## Problem 1: Linear Models
Read the dataset given in `r s_data_url` using the function `readr::read_csv()`. 


### Tasks

* Do a descriptive statistics on the given data
* Fit fixed linear models for `herd`, `sex` and both factors as fixed effects.



### Solution

Read the data into R

```{r}
tbl_data_beef <- readr::read_csv(file = s_data_url)
tbl_data_beef
```

### Descriptive Statistics for Weaning Weight

```{r}
summary(tbl_data_beef$`Weaning Weight`)
```

Summary is meaningful for continuous variables such as Weaning Weight. For discrete factor variables, the function table can be used

```{r}
table(tbl_data_beef$Herd)
```

The variable `sex` was not added to the dataset. If we had a variable that contains the sex of an animal, we could do the same summary with the table function as was done for herd.

### Fixed Linear Effect Model

```{r}
lm_ww_h <- lm(`Weaning Weight` ~ Herd, data = tbl_data_beef)
summary(lm_ww_h)
```

The above fits the herd as a regression coefficient, but what we want is a fixed effect model. We have to change the numbers in the herd column from numbers to factors. 

```{r}
tbl_data_beef$Herd <- as.factor(tbl_data_beef$Herd)
tbl_data_beef
```

The function class tells us the datatype of a column

```{r}
class(tbl_data_beef$Herd)
```

```{r}
lm_ww_h_fix <- lm(`Weaning Weight` ~ 0+ Herd, data = tbl_data_beef)
summary(lm_ww_h_fix)
```


## Additional Problem: Mixed Linear Effect Model
Start with the Sire model

$$y = X\beta + Zu + e$$

### Observations

```{r}
vec_y <- tbl_data_beef$`Weaning Weight`
vec_y
```


### Design Matrices


```{r}
mat_X <- matrix(data = c(1, 0,
                         1, 0,
                         1, 0,
                         1, 0,
                         0, 1,
                         0, 1,
                         0, 1,
                         0, 1,
                         1, 0,
                         1, 0,
                         1, 0,
                         1, 0,
                         0, 1,
                         0, 1,
                         0, 1,
                         0, 1), ncol = 2, byrow = TRUE)
mat_X
```


```{r}
tbl_data_beef$Herd
```

```{r}
mat_Z <- matrix(data = c(1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         1, 0, 0,
                         0, 1, 0,
                         0, 1, 0,
                         0, 1, 0,
                         0, 1, 0,
                         0, 1, 0,
                         0, 1, 0,
                         0, 0, 1,
                         0, 0, 1), ncol=3, byrow = TRUE)
mat_Z
```


```{r}
tbl_data_beef$Sire
```


## Solving MME

```{r}
lambda <- 1
mat_xtx <- t(mat_X) %*% mat_X
mat_xtx
```

```{r}
mat_xtx <- crossprod(mat_X)
mat_xtx
```

```{r}
mat_xtz <- crossprod(mat_X,mat_Z)
mat_xtz
```

```{r}
mat_ztx <- crossprod(mat_Z,mat_X)
mat_ztx <- t(mat_xtz)
mat_ztx
```

```{r}
mat_ztz <- crossprod(mat_Z)
mat_ztz
```

Adding $I * \lambda$ 

```{r}
mat_ilambda <- diag(1, nrow = nrow(mat_ztz), ncol = ncol(mat_ztz)) * lambda
mat_ilambda

```

```{r}
mat_ztzilambda <- mat_ztz + mat_ilambda
mat_ztzilambda
```


### Coefficient Matrix

```{r}
mat_coef <- rbind(cbind(mat_xtx, mat_xtz),
                  cbind(mat_ztx, mat_ztzilambda))
mat_coef
```


### Right Hand Side

```{r}
mat_xty <- crossprod(mat_X, vec_y)
mat_xty
```

```{r}
mat_zty <- crossprod(mat_Z, vec_y)
mat_zty
```

```{r}
mat_rhs <- rbind(mat_xty, mat_zty)
mat_rhs
```


### The Solution

```{r}
vec_unknown <- solve(mat_coef, mat_rhs)
vec_unknown
```

The first two components of the solution vector `vec_unknown` corresponds to estimates of the effects of the two herds. The remaining three components of vector ` vec_unknown` correspond to the three sire breeding values.
